% !TEX root = 0_main.tex
\chapter{Conclusion}\label{chap:conclusion}
We present \gls{tinygarble}, an automated tool that can generate optimized and compact circuits for Yao's \acrfull{gc} protocol.
We are the first to define the circuit generation for \acrshort{gc} as a logic synthesis problem and to leverage the powerful and established logic synthesis techniques with our custom objectives.
We improve the results of one of the best automated tools for \acrshort{gc} generation, \gls{pcf} \cite{kreuter2013pcf}, by several orders of magnitude. For instance, \gls{tinygarble} compacts the $\numprint{1024}$-bit multiplication by $\numprint{2504}$ times, while decreasing the number of non-XOR gates by 80\%; we compress the $\numprint{16000}$-bit Hamming distance by a factor of $\numprint{7345}$ times and with 47\% less non-XOR gates.
Further, \gls{tinygarble} can implement functions that have never been reported before, such as \acrshort{sha}-3.
We perform extensive benchmarking with both commercial and open source hardware synthesis tools and compare the results.

We introduce a gabled processor as a scalable solution for private function evaluation (\acrshort{pf-sfe}).
This garbled processor relies on the \gls{mips} architecture and the private function can be compiled using ubiquitous tools, e.g., gcc.
We also introduce \gls{arm2gc}, a simple-to-use and efficient garbled processor framework based on \gls{arm} and \gls{tinygarble}.
Users can develop secure functions in high-level languages and compile them using standard fully verified \gls{arm} cross-compilers.
Evaluations on a host of benchmark functions show that the \gls{arm2gc} framework achieves efficiency close to that of \gls{tinygarble} \acrshort{gc} synthesis.
The low cost of \gls{arm2gc} is made possible by \gls{skipgate}, a novel algorithm for the sequential \acrshort{gc}.
\gls{skipgate} omits the communication cost for gates with outputs independent of private data and also the gates not affecting the final output.
As a result of \gls{skipgate}, only the gates associated with private information in the massive \gls{arm} circuit incur communication cost.

Our \gls{tinygarble} and \gls{arm2gc} frameworks strongly improve the previous results and pave the way for a practical, scalable, and simple-to-use secure computation with many exciting applications.
For instance, \gls{tinygarble} is an enabling technology for performing \acrshort{gc} operations on mobile platforms, which is prohibitively expensive using the prior techniques.
And \gls{arm2gc} enables users with no or limited knowledge about the \acrshort{gc} protocol or Boolean circuit to develop privacy-preserving applications in high-level languages.
