% !TEX root = 0_main.tex
\chapter{Tinygarble GC Engine}\label{chap:engine}
In this section, we explain \gls{tinygarble} \acrshort{gc} Engine, an implementation of the Yao's \acrshort{gc} protocol for sequential circuits.
First, we describe the implementation of sequential \acrshort{gc} protocol in \gls{tinygarble} \acrshort{gc} engine.
Next, we describe \acrfull{sscd}, a format for storing and processing sequential Boolean circuits in our \acrshort{gc} engine and how \gls{verilog} \gls{netlist} is translated into \acrshort{sscd}.
We also explain the communication steps and pipelining of the \acrshort{gc} protocol for sequential circuits.
We then introduce the terminate signal that helps the user to identify when the circuit output is ready to reduce the garbling cost.
We implemented the \acrshort{gc} engine in C++, and its source code is available in the \gls{tinygarble} repository\footnote{\url{https://github.com/esonghori/\gls{tinygarble}}}.

\section{Yao's GC Protocol Implementation for Sequential Circuits} \label{sec:engine-gc}
\gls{tinygarble} \acrshort{gc} engine is an implementation of Yao's protocol for garbling and evaluation of sequential circuits.
We build it based on JustGarble's implementation of the \acrshort{gc} protocol \cite{bellare2013efficient}.
We choose JustGarble because it has a number of advantages over other available implementations: simplicity and efficiency (use of \acrshort{aes} block cipher and \acrshort{aes-ni} hardware accelerator).
However, JustGarble's realization comes with some shortcomings.
(i) It only realizes the garbling/evaluation procedures and does not include the communication capability between the parties.
Most importantly, it does not contain the implementation of \acrshort{ot} that is a crucial part of communication in the \acrshort{gc} protocol.
(ii) It does not include Half Gate, the recent optimization of the \acrshort{gc} protocol \cite{zahur2015two}.
Half Gate shrinks the \acrshort{gc} protocol cost by 33\% by reducing the communication required for an non-XOR gate from three labels to two.

Our \acrshort{gc} engine supports both garbling/evaluation and communication between the two parties including \acrshort{ot} (see \sect{sec:engine-comm}).
Our implementation of \acrshort{ot} also includes its extended version \cite{husted2013gpu} that reduced the computation cost by transferring the bulk of messages together.
We use OpenSSL library to handle big integer and modular exponentiation in \acrshort{ot} \cite{viega2002network}.
We also add Half Gate method to the implementation on top of other optimizations including Free XOR and Fixed-Key Block-Cipher.
Similar to JustGarble, we benefit from \acrshort{aes-ni}, a hardware implementation of \acrshort{aes} in Intel processors.

One of the prominent features of \gls{tinygarble} \acrshort{gc} engine is its support for garbling/evaluation of sequential circuits.
The gates in a sequential circuit are garbled/evaluated for multiple sequential cycles.
The parties agree on the total number of cycles before starting of the protocol.
To ensure security, the encryption tweaks $T$ (see \sect{sssec:prelim-aes}) for each gate has to be unique in each cycle\cite[Sect. 3.4]{henecka2013faster}.
We set tweak $T$ for each gate to be the concatenation of the cycle index (cid) and the unique gate identifiers (gid) in the combinational part of the sequential circuit, i.e., $T = \textrm{cid} || \textrm{gid}$.
An alternative method could be to use a monotonic counter in the garbling/evaluation routines that increases by one for each gate.
As in previous work, security and correctness of the \acrshort{gc} garbling/evaluation follow from the uniqueness of the tweak $T$ and the existing proofs of security and correctness are still valid for \gls{tinygarble} (see \cite{lindell2009proof, bellare2013efficient, zahur2015two}).


\section{Simple Sequential Circuit Description}\label{sec:engine-sscd}
JustGarble \cite{bellare2013efficient} developed the \acrfull{scd} format to represent combinational Boolean circuits.
We introduce a similar format called \acrfull{sscd} that additionally include information about the \acrshort{ff}s and the newly added terminate signal in sequential circuits.

Besides clock (clk) and reset (rst) ports that are not relevant to the \acrshort{gc} protocol, \acrshort{ff}s has three other ports: initial (I), input (D), output (Q).
The initial wires (I) are treated as inputs and can belong to either Alice or Bob.
We denote Alice's initial values as g\_init (g for Garbler) and Bob' as e\_init (e for Evaluator).
The input of the \acrshort{ff}s at any sequential cycle are fed by their outputs at the previous cycle.
The inputs of the \acrshort{ff}s (D) are treated  as the output of the circuit.
Similarly, the output of the \acrshort{ff}s (Q) are considered as the input of the circuit.
The Q wires at the first sequential cycles are set to their corresponding I's (provider by either Alice or Bob) and at the rest of cycles are set to their corresponding D's.
We denote Alice's and Bob's inputs as g\_input and e\_input respectively.

In the \acrshort{sscd} format, we index wires according to the following order: (1) g\_init (2) e\_init (3) g\_input (4) e\_input (5) gates' output.
\acrshort{sscd} is stored in a human-readable ASCII format.
An \acrshort{sscd} file consists of seven lines: (1) g\_init\_size, e\_init\_size, g\_input\_size, e\_input\_size, dff\_size, output\_size, terminate (an optional 1-bit output indicating that the circuit is done, see \sect{sec:engine-term}), and the number of gates. (2) gates' first input index. (3) gates' second input index. (4) gates' type. (5) outputs index. (6) Flip-Flops' D index. (7) Flip-Flops' I index (chosen from g\_init or e\_init).

\section{Translating Verilog Netlist to SSCD}\label{sec:engine-v2sscd}
The output of \gls{tinygarble} synthesis flow is an optimized gate level \gls{verilog} code called \gls{netlist}.
The \gls{netlist} format needs to be translated into \acrshort{sscd} format to be readable by \gls{tinygarble} \acrshort{gc} engine.
We developed a tool called V2SCD that converts a \gls{verilog} \gls{netlist} into an \acrshort{sscd} file.
V2SCD first parses the \gls{netlist} file and detect the gates and their ports.
Next, it topologically sorts the gates based on their dependencies through wire connections.
Finally, it creates a human-readable \acrshort{sscd} file.

\section{Communication}\label{sec:engine-comm}
JustGarble's implementation of the \acrshort{gc} protocol lacks the communication features required by the \acrshort{gc} protocol.
We implemented and integrated these features into our \acrshort{gc} engine.
The communication and computation in \gls{tinygarble} \acrshort{gc} engine follow the steps below:
\begin{enumerate}
\item Alice generates the garbled tables.
\item Alice sends her input labels to Bob.
\item Alice sends Bob's input labels to him through \acrshort{ot}.
\item Alice sends garbled tables to Bob.
\item Bob evaluates the circuit.
\item Alice sends the \acrfull{lsb} of labels of the output labels to Bob such that he can learn the output.
  We can change this to let Alice learn some or all bits of the output.
  To do so, Bob has to send the \acrshort{lsb} of the selected output labels to Alice.
\end{enumerate}

By default, the steps are the same for sequential circuits as well.
Alice has to generate the garbled circuit for $cc$ sequential cycles where $cc$ is a predetermined number set by the parties.
Alice sends labels corresponding to her initial values (g\_init) to Bob as well as the ones corresponding to her inputs (g\_input) for all $cc$ cycles.
The initial labels are sent only for the first cycle.
The same goes for Bob's initial and input values, but through \acrshort{ot}.

\subsection{Pipelining and Low Memory Mode}\label{sec:engine-memory}
In the simple communication scheme described above the memory required to store the labels is not scalable in terms of the number of sequential cycles $cc$.
Both Alice and Bob has to save all the labels in the memory for all $cc$ cycles.
To avoid such large memory footprint, we design a communication mode for sequential circuits that reduces the memory footprint by pipelining the garbling/evaluation.
Pipelining for the \acrshort{gc} protocol was first proposed in \cite{husted2013gpu} for combinational circuits where the circuit is divided into multiple chunks and communication is done between garbling each chuck.
However, the pipelining is more fit for sequential circuits where garbling of the circuit in each cycle creates a natural chunk for the pipeline.

The communication and computation in \gls{tinygarble} \acrshort{gc} engine for a sequential circuit in low memory mode follows the steps below:
(1) Alice generates garbled tables for one sequential cycle.
(2) If it is the first cycle, Alice sends her initial labels to Bob.
(3) If it is the first cycle, Alice sends Bob's initial labels to him through \acrshort{ot}.
(4) Alice sends her input labels for this cycle to Bob.
(5) Alice sends Bob's input labels for this cycle to him through \acrshort{ot}.
(6) Alice sends garbled tables for this cycle to Bob.
(7) Bob evaluates the circuit for one cycles.
(8) Go to step (9) if cycle reaches $cc$, if not go to (1)
(9) Alice and Bob exchange \acrshort{lsb} of output labels to learn the output.

Note that in this mode of communication, the number of invocation of \acrshort{ot} increases from 1 to $cc$.
Since we are using the \acrshort{ot} extension, the cost of one invocation of \acrshort{ot} is almost constant when the number of transferred labels is large enough (in our case larger than 128)\footnote{The cost of \acrshort{ot} is dominated by the computation of modular exponentiation. In \acrshort{ot} extension, the number of modular exponentiation is $\BigO{min(l, p)}$ where $l$ is the number of messages (labels) and $p$ is the security parameter.}.
This means that when the number of Bob's inputs (bit widths of e\_input times $cc$) is lager than 128, the cost of transferring labels through \acrshort{ot} increases using pipelining.
Therefore, there is a trade-off between the cost of multiple invocations of \acrshort{ot} and the memory footprint when the number of Bob's input or sequential cycles is large.

\section{Terminate Signal}\label{sec:engine-term}
As explained in \chap{chap:seq}, sequential circuits are required to be garbled/evaluated for a number of sequential cycles ($cc$ cycles).
$cc$ has to be determined regardless of the inputs to ensure the privacy.
To do so, it is set such that the circuit functionality is ensured to be finished during $cc$ cycles, i.e., the worst case scenario.
There are sequential circuits that, given specific inputs, can complete the computation faster than their predetermined number of cycles.
For example, the sequential circuit presented in \cite{riazi2017toward} for Stable Match algorithm of $n$ pairs can finish the computation in $\BigO{n}$ while in the worst case scenario, the number of cycles is $\BigO{n^2}$.

To reduce the unnecessary cost after the computation is finished, \gls{tinygarble} enables users to add a one-bit output to a sequential circuit that indicates in which cycle the output of the circuit is ready.
We call this output the \textit{terminate signal}.
The use of the terminate signal was first proposed in \cite{riazi2017toward} for a particular application of stable matching using sequential \acrshort{gc}.
We generalized this approach for all sequential circuits.
e terminate signal enables users to stop garbling/evaluation of the circuit when no further computation is needed, avoiding futile communication and computation.
In \gls{tinygarble}, the signal is revealed every $K$ clock cycles where $1 \le K \le cc$.
The parties determine $k$ before starting the computation.
Depending on the structure of the sequential circuit and its functionality, revealing the terminate signal leaks some information about the inputs.
Thus, it has to be used with careful consideration.
$K=1$ shows the exact number of cycles for the given inputs to finish the function.
Larger $K$ exposes less information and $K=cc$ reveals no information about the inputs.
Therefore, revealing the terminate signal offers a trade-off between the privacy and the cost of garbling a sequential circuit.

\section{SkipGate Implementation}\label{sec:engine-skipgate}
\gls{tinygarble} \acrshort{gc} engine supports the \gls{skipgate} algorithm on top of its implementation of the \acrshort{gc} protocol.
We change the code of the engine according to the \gls{skipgate} algorithm described in \chap{chap:skipgate}.
Two new ports are allowed in the circuit for public input variables: (i) Public initial value (p\_init) that is connected to I input of \acrshort{ff}s and is read only at the first cycle. (ii) Public input (p\_input) that is directly connected to the gates of the circuit and is read every sequential cycle.
The first line of the \acrshort{sscd} format also includes the size of these two new ports: p\_init\_size,  g\_init\_size, e\_init\_size, p\_input\_size, g\_input\_size, e\_input\_size, dff\_size, output\_size, terminate\_id, and gate\_size.
