% !TEX root = 0_main.tex
\chapter*{Acknowledgements}
\thispagestyle{empty}

The completion of this thesis would not have been possible without the contribution of several individuals who in one way or another have provided their invaluable guidance and assistance.

My deepest gratitude to my advisor Professor Farinaz Koushanfar for her invaluable support, encouragement, and advice. Her dedication, fondness, and motivation towards doing novel research, as well as exceptional care for her students have taught me valuable lessons and greatly influenced my life at academic and especially personal levels.

I would also like to thank Professors Joseph R. Cavallaro, Aydin Babakhani, and Dan Wallach for serving as my committee members even at hardship. I also want to thank them for their brilliant comments and suggestions on the thesis. Special thanks to Professors Ahmad-Reza Sadeghi and Thomas Schneider of TU Darmstadt for their collaboration with our lab at Rice and later UCSD and their great academic and scientific supports.

I would also like to thank my collages: Siam U. Hussain, M. Sadegh Riazi, Ghada Dessouky, and Shaza Zeitouni for their endless help; this work would not have been possible without your assistance, seriously. Thank you!

Special thanks to my family. Words cannot express how grateful I am to my mother, father, and two brothers for all of the sacrifices that you have made on my behalf. In the end, I would like express appreciation to my beloved fianc\'ee, Zeinab, who was always my support in the moments when there was no one to answer my queries.

\clearpage
